---
title: "Andy's Modeling Updates"
author: "Andy Shen"
date: "`r Sys.Date()`"
header-includes:
  - \usepackage{booktabs}
output: 
  beamer_presentation:
    theme: "Warsaw"
    colortheme: "seahorse"
    slide_level: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      fig.align = "center")
library(tidyverse)
library(kableExtra)
library(htmltools)
library(xtable)
library(nnet)
```



## Agenda 4/6/2023

- Discuss R Project workflow and data pre-processing script

- Andy's updates on modeling

- "Evaluating situational decomposition" qualms

- Discuss structure of report

## R Projects

- "Sandbox" workspace for a specific project

- Main advantage is there is no need to change working directories or paths

- There is only 1 R project: `code`

- To open the project, double click `code.Rproj` (should open a new RStudio window)

- You must have open the `.Rproj` file before opening and running any `.R` or `.Rmd` file

## Data pre-processing

- There is an R script in the `code/` directory

- Judgment calls and decision rules

```{r}
# reading in data and recoding it
df_first_study <- read_csv("../processed-data/intox_sex_file.csv")

df_first_study$sex <- as.character(df_first_study$sex)

df_first_study$intoxication <- factor(df_first_study$intoxication)
df_first_study$intoxication <- relevel(df_first_study$intoxication, ref = "never")

df_first_study$intercourse_dummy <- factor(df_first_study$intercourse_dummy)

df_first_study$intox_most_recent_sex_words <- factor(df_first_study$intox_most_recent_sex_words)
df_first_study$intox_most_recent_sex_words <- relevel(df_first_study$intox_most_recent_sex_words, ref = "never")


df_1st_logistic <- df_first_study

# creating dummy variables for intoxication in case we need it later.
# df_1st_logistic <- df_first_study %>% 
#   dplyr::mutate(
#     never_intox = ifelse(intoxication == "never", 1, 0),
#     occasion_intox = ifelse(intoxication == "occasionally", 1, 0),
#     frequently_intox = ifelse(intoxication == "frequently", 1, 0)
#   )
```


```{r, eval=TRUE}
# running inividual logistic regressions
tassoc_1st <- glm(intercourse_dummy ~ intoxication, 
                  data = df_1st_logistic, 
                  family = "binomial")
# summary(tassoc_1st)
odds_ratio_mf <- exp(tassoc_1st$coefficients) #%>% log()

tassoc_male_1st <- glm(intercourse_dummy ~ intoxication, 
                  data = df_1st_logistic %>% dplyr::filter(sex == 1), 
                  family = "binomial")
# summary(tassoc_male_1st)
odds_ratio_m <- exp(tassoc_male_1st$coefficients) #%>% log()

tassoc_female_1st <- glm(intercourse_dummy ~ intoxication, 
                  data = df_1st_logistic %>% dplyr::filter(sex == 2), 
                  family = "binomial")
# summary(tassoc_female_1st)
odds_ratio_f <- exp(tassoc_female_1st$coefficients) #%>% log()
```

# Replicating Felson Study 1: Effect of intoxication on sexual intercourse

## Judgment calls

*Note:* all of the code can be found in the accompanying `.Rmd` file for these slides

- Removing NAs in outcome (intercourse)

- People who refused to respond were categorized in the reference category (no for sex, never for alcohol)

## Total association with logistic regression

```{r}
# inputting our numbers into a table for display
df_1st_OR <- data.frame(
  all_gender = odds_ratio_mf[-1] %>% rev(),
  male = odds_ratio_m[-1] %>% rev(),
  female = odds_ratio_f[-1] %>% rev() # do not need intercept coef
) %>% t() %>% data.frame() %>% tibble()
rownames(df_1st_OR) <- c("all_gender", "male", "female")
names(df_1st_OR) <- c("Occasionally", "Frequently")

df_1st_OR_kbl <- df_1st_OR %>% 
  rownames_to_column("Gender") %>% 
  dplyr::mutate(OR_diff = abs(Occasionally - Frequently)) %>% 
  kable(format = "latex", booktabs = TRUE, digits = 1,
      caption = "Total association logistic regression odds ratios (Andy).") %>%
  kableExtra::kable_styling(font_size = 10)
df_1st_OR_kbl
```

```{r}
# inputting felson's numbers
df_1st_Felson <- data.frame(
  all_gender = c(3.98, 8.48),
  male = c(3.57, 8.69),
  female = c(4.41, 7.69)
) %>% t() %>% data.frame() %>% tibble()
rownames(df_1st_Felson) <- c("all_gender", "male", "female")
names(df_1st_Felson) <- c("Occasionally", "Frequently")

# display purposes
df_1st_Felson_kbl <- df_1st_Felson %>% 
  rownames_to_column("Gender") %>% 
  dplyr::mutate(OR_diff = abs(Occasionally - Frequently)) %>% 
  kable(format = "latex", booktabs = TRUE, digits = 1,
      caption = "Total association logistic regression odds ratios (Felson et al.).") %>% 
  kableExtra::kable_styling(font_size = 10)
df_1st_Felson_kbl
```

## Takeaways

- Let's assume this subsample is representative of the larger sample used in the paper

- In this case, our analysis has "closed the gap" in terms of odds ratios

- Perhaps the tendency to have sex isn't *that* drastic for occasional vs frequent drinkers?

- How does this relate to spuriousness?


## Spuriousness of intoxication on sober sex

```{r}
spurious_assoc_1st <- nnet::multinom(intox_most_recent_sex_words ~ intoxication,
                                     data = df_1st_logistic, trace = FALSE)
# summary(spurious_assoc_1st)
```

```{r}
spur_coefs_all <- coef(spurious_assoc_1st)[, -1][-1, ] %>% rev()
tot_coefs_all <- tassoc_1st$coefficients[-1] %>% rev()
# spur_coefs_all; tot_coefs_all
spuriousness_all <- spur_coefs_all / tot_coefs_all * 100
```

```{r}
spurious_assoc_male_1st <- nnet::multinom(intox_most_recent_sex_words ~ intoxication,
                                     data = df_1st_logistic %>% dplyr::filter(sex == 1),
                                     trace = FALSE)
```

```{r}
spur_coefs_male <- coef(spurious_assoc_male_1st)[, -1][-1, ] %>% rev()
tot_coefs_male <- tassoc_male_1st$coefficients[-1] %>% rev()
spuriousness_male <- spur_coefs_male / tot_coefs_male * 100
```


```{r}
spurious_assoc_female_1st <- nnet::multinom(intox_most_recent_sex_words ~ intoxication,
                                     data = df_1st_logistic %>% dplyr::filter(sex == 2),
                                     trace = FALSE)
```

```{r}
spur_coefs_female <- coef(spurious_assoc_female_1st)[, -1][-1, ] %>% rev()
tot_coefs_female <- tassoc_female_1st$coefficients[-1] %>% rev()
spuriousness_female <- spur_coefs_female / tot_coefs_female * 100
```




```{r}
spur_df <- data.frame(
  "all_gender"  = spuriousness_all,
  "males" = spuriousness_male,
  "females" = spuriousness_female
) %>% t() %>% data.frame()
names(spur_df) <- c("Occasionally", "Frequently")
spur_df %>% 
  knitr::kable(digits = 1, 
               booktabs = TRUE,
               format = 'latex',
               caption = "Our spuriousness values") %>% 
  kableExtra::kable_styling(font_size = 10)
```

```{r}
felson_1st_spur <- data.frame(
  "Occasionally" = c(95.7, 95.3, 97.3),
  "Frequently" = c(91.6, 91.2, 93.6)
)
rownames(felson_1st_spur) <- c("all_gender", "males", "females")
felson_1st_spur %>% 
  knitr::kable(booktabs = TRUE,
               format = 'latex',
               caption = "Felson's spuriousness values") %>% 
  kableExtra::kable_styling(font_size = 10)
```



# Replicating Felson Study 2: Effect of intoxication on contraceptive use

```{r}
df_second_study <- read_csv("../processed-data/intox_contraception_file.csv") %>% 
  dplyr::select(-intox_most_recent_sex, -intercourse_dummy, -contraception_use, -intox_most_recent_sex_words) %>% 
  dplyr::mutate_at(vars(sex, intoxication, contraception_use_words, contraception_sex), factor)
```


## Data pre-processing

```{r}
df_second_study$contraception_sex <- relevel(df_second_study$contraception_sex, ref = "use")
df_second_study$contraception_use_words <- relevel(df_second_study$contraception_use_words, ref = "use")
df_second_study$intoxication <- relevel(df_second_study$intoxication, ref = "never")

df_2nd_logistic <- df_second_study
```


- Same R script: `data-preprocessing.R`

- Reduce cases down to respondents who have had sex ($n=2565$)

## Judgment Calls

- NAs in contraception use were not given benefit of the doubt

## Total association using logistic regression

```{r, eval=TRUE}
# running inividual logistic regressions
tassoc_2nd <- glm(contraception_use_words ~ intoxication, 
                  data = df_2nd_logistic, 
                  family = "binomial")
# summary(tassoc_2nd)
odds_ratio_mf <- exp(tassoc_2nd$coefficients) #%>% log()

tassoc_male_2nd <- glm(contraception_use_words ~ intoxication, 
                  data = df_2nd_logistic %>% dplyr::filter(sex == 1), 
                  family = "binomial")
# summary(tassoc_male_2nd)
odds_ratio_m <- exp(tassoc_male_2nd$coefficients) #%>% log()

tassoc_female_2nd <- glm(contraception_use_words ~ intoxication, 
                  data = df_2nd_logistic %>% dplyr::filter(sex == 2), 
                  family = "binomial")
# summary(tassoc_female_2nd)
odds_ratio_f <- exp(tassoc_female_2nd$coefficients) #%>% log()
```

```{r}
# inputting our numbers into a table for display
df_2nd_OR <- data.frame(
  all_gender = odds_ratio_mf[-1] %>% rev(),
  male = odds_ratio_m[-1] %>% rev(),
  female = odds_ratio_f[-1] %>% rev() # do not need intercept coef
) %>% t() %>% data.frame() %>% tibble()
rownames(df_2nd_OR) <- c("all_gender", "male", "female")
names(df_2nd_OR) <- c("Occasionally", "Frequently")

df_2nd_OR_kbl <- df_2nd_OR %>% 
  rownames_to_column("Gender") %>% 
  dplyr::mutate(OR_diff = abs(Occasionally - Frequently)) %>% 
  kable(format = "latex", booktabs = TRUE, digits = 2,
      caption = "Total association logistic regression odds ratios (Andy).") %>%
  kableExtra::kable_styling(font_size = 10)
df_2nd_OR_kbl
```

```{r}
# inputting felson's numbers
df_2nd_Felson <- data.frame(
  all_gender = c(1.05, 1.38),
  male = c(1.00, 1.40),
  female = c(1.07, 1.51)
) %>% t() %>% data.frame() %>% tibble()
rownames(df_2nd_Felson) <- c("all_gender", "male", "female")
names(df_2nd_Felson) <- c("Occasionally", "Frequently")

# display purposes
df_2nd_Felson_kbl <- df_2nd_Felson %>% 
  rownames_to_column("Gender") %>% 
  dplyr::mutate(OR_diff = abs(Occasionally - Frequently)) %>% 
  kable(format = "latex", booktabs = TRUE, digits = 2,
      caption = "Total association logistic regression odds ratios (Felson et al.).") %>% 
  kableExtra::kable_styling(font_size = 10)
df_2nd_Felson_kbl
```


## Takeaways

- This seems pretty good

- Now let's discuss spuriousness



### Summary

- x

- y

- z


## Spuriousness of intoxication on contraception use

```{r}
spurious_assoc_2nd <- nnet::multinom(contraception_sex ~ intoxication,
                                     data = df_2nd_logistic, trace = FALSE)
# summary(spurious_assoc_2nd)
```

```{r}
spur_coefs_all <- coef(spurious_assoc_2nd)[, -1][-1, ] %>% rev()
tot_coefs_all <- tassoc_2nd$coefficients[-1] %>% rev()
# spur_coefs_all; tot_coefs_all
spuriousness_all <- spur_coefs_all / tot_coefs_all * 100
```

```{r}
spurious_assoc_male_2nd <- nnet::multinom(contraception_sex ~ intoxication,
                                     data = df_2nd_logistic %>% dplyr::filter(sex == 1),
                                     trace = FALSE)
```

```{r}
spur_coefs_male <- coef(spurious_assoc_male_2nd)[, -1][-1, ] %>% rev()
tot_coefs_male <- tassoc_male_2nd$coefficients[-1] %>% rev()
spuriousness_male <- spur_coefs_male / tot_coefs_male * 100
```


```{r}
spurious_assoc_female_2nd <- nnet::multinom(contraception_sex ~ intoxication,
                                     data = df_2nd_logistic %>% dplyr::filter(sex == 2),
                                     trace = FALSE)
```

```{r}
spur_coefs_female <- coef(spurious_assoc_female_2nd)[, -1][-1, ] %>% rev()
tot_coefs_female <- tassoc_female_2nd$coefficients[-1] %>% rev()
spuriousness_female <- spur_coefs_female / tot_coefs_female * 100
```




```{r}
spur_df <- data.frame(
  "all_gender"  = spuriousness_all,
  "males" = spuriousness_male,
  "females" = spuriousness_female
) %>% t() %>% data.frame()
names(spur_df) <- c("Occasionally", "Frequently")
spur_df %>% 
  knitr::kable(digits = 1, 
               booktabs = TRUE,
               format = 'latex',
               caption = "Our spuriousness values") %>% 
  kableExtra::kable_styling(font_size = 10)
```

```{r}
felson_2nd_spur <- data.frame(
  "Occasionally" = rep(NA, 3),
  "Frequently" = c(46.9, 41.1, 68.3)
)
rownames(felson_2nd_spur) <- c("all_gender", "males", "females")
felson_2nd_spur %>% 
  knitr::kable(booktabs = TRUE,
               format = 'latex',
               caption = "Felson's spuriousness values") %>% 
  kableExtra::kable_styling(font_size = 10)
```

## Side note

- Both the coefficients from binary LR and multinomial LR are significant for frequent female drinkers only in Felson's study. 

- This happens to be the only result with a reasonable spuriousness value!


## Issues with situational decomposition

- What happens when there is above 100% spuriousness?

- What happens when the coefficient is negative? (Felson seems to wave it away)


## Next steps

- Andy 

  - Write up today's slides into a latex document called `final-report.tex` (formatted by TD)
  
  - Start preprocessing data for instrumental variables analysis
  
  - Look into doing IVLS with binary outcome (IV logistic regression)

- Tiffany

  - 

