---
title: "Andy's Modeling Updates"
author: "Andy Shen"
date: "`r Sys.Date()`"
header-includes:
  - \usepackage{booktabs}
output: 
  beamer_presentation:
    theme: "Warsaw"
    colortheme: "seahorse"
    slide_level: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      fig.align = "center")
library(tidyverse)
library(kableExtra)
library(htmltools)
library(xtable)
```



## Agenda 4/6/2023

- Discuss R Project workflow and data pre-processing script

- Andy's updates on modeling

- "Evaluating situational decomposition" qualms

- Discuss structure of report

## R Projects

- "Sandbox" workspace for a specific project

- Main advantage is there is no need to change working directories or paths

- There is only 1 R project: `code`

- To open the project, double click `code.Rproj` (should open a new RStudio window)

- You must have open the `.Rproj` file before opening and running any `.R` or `.Rmd` file

## Data pre-processing

- There is an R script in the `code/` directory

```{r}
# reading in data and recoding it
df_first_study <- read_csv("../processed-data/intox_sex_file.csv")

df_first_study$sex <- as.character(df_first_study$sex)

df_first_study$intoxication <- factor(df_first_study$intoxication)
df_first_study$intoxication <- relevel(df_first_study$intoxication, ref = "never")

df_first_study$intercourse_dummy <- factor(df_first_study$intercourse_dummy)

# creating dummy varaibles for intoxication in case we need it later.
df_1st_logistic <- df_first_study %>% 
  dplyr::mutate(
    never_intox = ifelse(intoxication == "never", 1, 0),
    occasion_intox = ifelse(intoxication == "occasionally", 1, 0),
    frequently_intox = ifelse(intoxication == "frequently", 1, 0)
  )
```


```{r, eval=TRUE}
# running inividual logistic regressions
tassoc_1st <- glm(intercourse_dummy ~ intoxication, 
                  data = df_1st_logistic, 
                  family = "binomial")
# summary(tassoc_1st)
odds_ratio_mf <- exp(tassoc_1st$coefficients)

tassoc_male_1st <- glm(intercourse_dummy ~ intoxication, 
                  data = df_1st_logistic %>% dplyr::filter(sex == 1), 
                  family = "binomial")
# summary(tassoc_male_1st)
odds_ratio_m <- exp(tassoc_male_1st$coefficients)

tassoc_female_1st <- glm(intercourse_dummy ~ intoxication, 
                  data = df_1st_logistic %>% dplyr::filter(sex == 2), 
                  family = "binomial")
# summary(tassoc_female_1st)
odds_ratio_f <- exp(tassoc_female_1st$coefficients)
```

# Replicating Felson Study 1: Effect of intoxication on sexual intercourse

## Judgment calls

*Note:* all of the code can be found in the accompanying `.Rmd` file for these slides

- Removing NAs in outcome (intercourse)

- People who refused to respond were categorized in the reference category (no for sex, never for alcohol)

## Total Association with logistic regression

```{r}
df_1st_OR <- data.frame(
  all_gender = odds_ratio_mf[-1] %>% rev(),
  male = odds_ratio_m[-1] %>% rev(),
  female = odds_ratio_f[-1] %>% rev() # do not need intercept coef
) %>% t() %>% data.frame() %>% tibble()
rownames(df_1st_OR) <- c("all_gender", "male", "female")
names(df_1st_OR) <- c("Occasionally", "Frequently")
df_1st_OR_kbl <- df_1st_OR %>% 
  rownames_to_column("Gender") %>% 
  dplyr::mutate(OR_diff = abs(Occasionally - Frequently)) %>% 
  kable(format = "latex", booktabs = TRUE, digits = 1,
      caption = "Total association logistic regression odds ratios (Andy).") %>%
  kableExtra::kable_styling(font_size = 10)
df_1st_OR_kbl
```

```{r}
df_1st_Felson <- data.frame(
  all_gender = c(3.98, 8.48),
  male = c(3.57, 8.69),
  female = c(4.41, 7.69)
) %>% t() %>% data.frame() %>% tibble()
rownames(df_1st_Felson) <- c("all_gender", "male", "female")
names(df_1st_Felson) <- c("Occasionally", "Frequently")
df_1st_Felson_kbl <- df_1st_Felson %>% 
  rownames_to_column("Gender") %>% 
  dplyr::mutate(OR_diff = abs(Occasionally - Frequently)) %>% 
  kable(format = "latex", booktabs = TRUE, digits = 1,
      caption = "Total association logistic regression odds ratios (Felson et al.).") %>% 
  kableExtra::kable_styling(font_size = 10)
df_1st_Felson_kbl
```


## Spurious effect using multinomial logistic regression


